\newglossary[slg]{symbolslist}{syi}{syg}{Symbols}
% ==== MAIN GLOSSARY ====
\newglossaryentry{conjecture}{
	name={conjecture},
	description={what mathematicians call their guesses, similarly to how scientists call their guesses hypotheses.}
}

\newglossaryentry{theorem}{
	name={theorem},
	description={conjectures that have been proven.}
}

\newglossaryentry{deductive reasoning}{
	name={deductive reasoning},
	description={the foundation on which proofs are based. Proofs play a central role in mathematics.}
}

\newglossaryentry{argument}{
	name={argument},
	description={a set of premises and a conclusion. An argument is valid if it is the case that, if the premises are true then the conclusion must be true.}
}

\newglossaryentry{premise}{
	name={premise},
	description={a fact or assumption in an argument.}
}

\newglossaryentry{conclusion}{
	name={conclusion},
	description={the end product of an argument. In a valid argument the conclusion will be true if the given premises are true.}
}

\newglossaryentry{connective symbol}{
	name={connective symbol},
	description={stand-ins for some of the words used to combine statements in an argument.}	
}

\newglossaryentry{conjunction}{
	name={conjunction},
	description={P \gls{symb:and} Q is the conjunction of P and Q.}
}

\newglossaryentry{disjunction}{
	name={disjunction},
	description={P \gls{symb:or} Q is the disjunction of P and Q.}
}

\newglossaryentry{negation}{
	name={negation},
	description={\gls{symb:not} P is the negation of P.}
}

\newglossaryentry{well-formed formula}{
	name={well-formed formula},
	description={a "grammatical" expression in the language of logic.}
}


% ==== SYMB GLOSSARY ====
\newglossaryentry{symb:therefore}{
	name=$\therefore$,
	description={therefore, which is used for an argument's conclusion.},
	type=symbolslist
}
\newglossaryentry{symb:not}{
	name=$\lnot$,
	description={a logical negation. Note that this symbol always applies only to the statement that comes immediately after it.},
	type=symbolslist
}
\newglossaryentry{symb:and}{
	name=$\land$,
	description={a logical AND.},
	type=symbolslist
}
\newglossaryentry{symb:or}{
	name=$\lor$,
	description={a logical OR. Note that this is the inclusive or.},
	type=symbolslist
}

