\documentclass{article}
\input{preamble}

\title{Proofs}
\author{Axel Sorenson}
\date{July 11th, 2024}

\begin{document}
\maketitle

\section{Proof Strategies}
Mathematicians usually state the answer to a mathematical question in the form of a \gls{theorem}, such that if the assumptions stated within (which are called the \glspl{hypothesis}) are true, then some conclusion must also be true. Free variables within the theorem, if assigned specific values from the \gls{universe of discourse}, turns the theorem into an \gls{instance} of the theorem. For all instances of the theorem such that the hypotheses are true, the conclusion must also be true in order for the theorem to be correct. If even one instance has all true hypotheses and a false conclusion, then the theorem is incorrect. This is called a \gls{counterexample} to the theorem.\\

\noindent The following is a general structure of a proof:\\
\begin{tikzpicture}[node distance=1cm, auto]
    % Define the boxes
    \node[draw, rectangle, text width=5cm, align=center] (box1) {Hypothesis 1, Hypothesis 2, Hypothesis 3, ...};
    \node[draw, rectangle, text width=5cm, align=center, right=of box1] (box2) {\textit{Assuming} Hypothesis 1 is true, then we can \textit{assert} that Assumption 1 is true. Assumption 1 can now be treated exactly as if it was one of the given hypotheses, allowing us to assert other truths from it.};
    \node[draw, rectangle, text width=5cm, align=center, below=of box2] (box3) {Given that Assumption 1 is true, then we can \textit{assert} that Assumption 2 is true.};
    \node[draw, rectangle, text width=5cm, align=center, left=of box3] (box4) {...};
    \node[draw, rectangle, text width=5cm, align=center, below=of box4] (box5) {Given Assumption ... is true, we can \textit{assert} that Conclusion is true.};
    \node[draw, rectangle, text width=5cm, align=center, right=of box5] (box6) {Therefore, Conclusion is true.};
    
    % Draw arrows between the boxes
    \draw[->] (box1) -- (box2);
    \draw[->] (box2) -- (box3);
    \draw[->] (box3) -- (box4);
    \draw[->] (box4) -- (box5);
    \draw[->] (box5) -- (box6);
\end{tikzpicture} \\
We can also transform proofs into equivalent proofs that are easier to write about, much like how we can transform equations (such as adding or subtracting values on both sides of an equation). So we can do a transformation of a simple proof like this:

\begin{tikzpicture}[node distance=1cm, auto]
    \node[draw, rectangle, text width=5cm, align=center] (box1) {\Gls{theorem}: If $P$ is true, then $Q$ is true (The conclusion that we must prove to be true is $P \text{ \gls{symb:implies} } Q$)};
    \node[draw, rectangle, text width=5cm, align=center, right=of box1] (box2) {Assume $P$ is true, making it equivalent to a hypothesis. The conclusion would then be that $Q$ is true.};
    \node[draw, rectangle, text width=5cm, align=center, below=of box2] (box3) {We must now show that if $P$ is true, then $Q$ must be true in all cases.};
    
    \draw[->] (box1) -- (box2);
    \draw[->] (box2) -- (box3);
\end{tikzpicture} \\
Form of the proof:\\
Suppose $P$. [Proof of $Q$ goes here]. Therefore $P \rightarrow Q$.\\

\noindent This transformation essentially means that if our theorem we are trying to prove has the form $P \rightarrow Q$, then we can assume $P$ (the \gls{antecedent}) to be true, adding it to our list of hypotheses, and then change the conclusion from $P \rightarrow Q$ to $Q$ (the \gls{consequent}). This new form of the proof might be easier to solve, depending on the complexity of what $P$ and $Q$ is. Note that this doesn't solve the proof - it is just a transformation that can be used to create a new (hopefully easier) problem that must be solved.\\

\noindent A proof may have several transformations applied to it over time, and it's important that we keep track of the results of this sequence of transformations. We therefore introduce the following terminology. We will refer to the statements that are known or assumed to be true at some point in the course of figuring out a proof as \glspl{given}, and the statement that remains to be proven at that point as the \gls{goal}. When you are starting to figure out a proof, the givens will be just the hypotheses of the theorem you are proving, but they may later include other statements that have been inferred from the hypotheses or added as new assumptions as the result of some transformation of the problem. The goal will initially be the conclusion of the theorem, but it may be changed several times in the course of figuring out a proof.\\

\noindent Note that the transformation technique described earlier is best described as enacted upon givens and goals as it can be used on transformations of the original proof rather than just on the proof itself. Here's an example of this transformation in action:

\noindent Suppose $a$ and $b$ are real numbers. Prove that if $0 < a < b$ then $a^2 < b^2$.\\
Transformation:\\
\begin{tikzpicture}[node distance=1cm, auto]
    \node[draw, rectangle, text width=5cm, align=center] (box1) {
        Givens:\\
        $a$ and $b$ are real numbers
    };
    \node[draw, rectangle, text width=5cm, align=center, right=of box1] (box2) {
        Goal:\\
        $(0 < a < b) \rightarrow (a^2 < b^2)$
    };

    \node[draw, rectangle, text width=5cm, align=center, below=of box1] (box3) {
        Givens:\\
        $a$ and $b$ are real numbers\\
        $(0 < a < b)$
    };
    
    \node[draw, rectangle, text width=5cm, align=center, below=of box2] (box4) {
        Goal:\\
        $(a^2 < b^2)$
    };

    \draw[->] (box1) -- (box3);
    \draw[->] (box2) -- (box4);
\end{tikzpicture} \\
Multiplying $a < b$ by $a$ gives us $a^2 < ab$ and multiplying $a < b$ by $b$ gives us $ab < b^2$. Together that makes $a^2 < ab < b^2$, making our goal, $a^2 < b^2$, true.\\

\noindent Transformations like these, when translated from the scratchwork of a proof to a formalized proof designed for other readers, often involve one or two sentences at the beginning and end of the proof to communicate this transformation. The work in the middle in between the beginning and end will then involve solving this new transformation of the problem. 

\section{Proofs Involving Negations and Conditionals}
When your goal is in the form $\neg P$, you should try to transform the problem so that the goal is not negated. If this isn't possible, then it's usually best to do a proof by contradiction.\\
\begin{tikzpicture}[node distance=1cm, auto]
    \node[draw, rectangle, text width=5cm, align=center] (box1) {
        Givens:\\
        \textemdash
    };
    \node[draw, rectangle, text width=5cm, align=center, right=of box1] (box2) {
        Goal:\\
        $\neg P$
    };

    \node[draw, rectangle, text width=5cm, align=center, below=of box1] (box3) {
        Givens:\\
        $P$
    };
    
    \node[draw, rectangle, text width=5cm, align=center, below=of box2] (box4) {
        Goal:\\
        $Contradiction$
    };

    \draw[->] (box1) -- (box3);
    \draw[->] (box2) -- (box4);
\end{tikzpicture} \\
Form of the proof:\\
Suppose $P$ is true. [Proof of contradiction goes here]. Thus, $P$ is false.\\

\noindent Note, that proof of contradictions often involve contradicting one of your givens (you can even try contradicting the given we just added, $P$).\\

\noindent Proving a goal by contradiction has the advantage that it allows you to assume that your conclusion is false, providing you with another given to work with.

\subsection{Exercises}
1a. Suppose $P \rightarrow Q$ and $Q \rightarrow R$ are both true. Prove that $P \rightarrow R$ is true.\\

\begin{tikzpicture}
    \beginProofBoxes{
        $P \rightarrow Q$\\
        $Q \rightarrow R$\\
    }{
        $P \rightarrow R$
    }

    \addProofBoxes{
        $P \rightarrow Q$\\
        $Q \rightarrow R$\\
        $P$
    }{
        $R$
    }

    \addProofBoxes{
        $P \rightarrow Q$\\
        $Q \rightarrow R$\\
        $P$\\
        $Q$\\
    }{
        $R$
    }

\end{tikzpicture}\\
Formalized proof:\\
Suppose $P$. Given $P$ and $P \rightarrow Q$, we know $Q$ is true. Given $Q$ and $Q \rightarrow R$, we know $R$ is true. Thus, if $P$ is true, then $R$ is true.\\

\noindent 1b. Suppose $\neg R \rightarrow (P \rightarrow \neg Q)$ is true. Prove that $P \rightarrow (Q \rightarrow R)$ is true.\\

\begin{tikzpicture}
    \beginProofBoxes{
        $\neg R \rightarrow (P \rightarrow \neg Q)$
    }{
        $P \rightarrow (Q \rightarrow R)$
    }

    \addProofBoxes{
        $\neg R \rightarrow (P \rightarrow \neg Q)$\\
        $P$
    }{
        $Q \rightarrow R$
    }

    \addProofBoxes{
        $\neg R \rightarrow (P \rightarrow \neg Q)$\\
        $P$\\
        $Q$
    }{
        $R$
    }

    \addProofBoxes{
        $\neg R \rightarrow (P \rightarrow \neg Q)$\\
        $P$\\
        $Q$\\
        $\neg R$
    }{
        \textit{Contradiction}
    }

    \addProofBoxes{
        $\neg R \rightarrow (P \rightarrow \neg Q)$\\
        $P$\\
        $Q$\\
        $\neg R$\\
        $P \rightarrow \neg Q$
    }{
        \textit{Contradiction}
    }
\end{tikzpicture}\\
Formalized Proof:\\
Suppose $P$. Suppose $Q$. Suppose $\neg R$. Given $\neg R$ and $\neg R \rightarrow (P \rightarrow \neg Q)$, we know $P \rightarrow \neg Q$ is true. Given $P$ and $P \rightarrow \neg Q$, we know $\neg Q$ is true. However, $Q$ is also true. We have hit a contradiction. Thus, $R$ is true. Given that $R$ is true when $Q$ is true, then $Q \rightarrow R$ is true. Given that $Q \rightarrow R$ is true when $P$ is true, then $P \rightarrow (Q \rightarrow R)$ is true. 

\section{Proofs Involving Quantifiers}
If your proof makes no assumptions to what $x$ is, then we say that $x$ is \gls{arbitrary}. In particular, you must not assume that $x$ is equal to any other object already under discussion in the proof. Thus, if the letter $x$ is already being used in the proof to stand for some particular object, then you cannot use it to stand for an arbitrary object. So, proving $\forall xP(x)$ involves ensuring that $x$ is arbitrary, and that a proof for $\forall xP(x)$ holds for all possible values $x$ can be.\\

\noindent Strategy:\\
\begin{tikzpicture}
    \beginProofBoxes{
        \textemdash
    }{
        $\forall xP(x)$
    }

    \addProofBoxes{
        \textemdash
    }{
        $P(x)$
    }
\end{tikzpicture}\\
Formalized Proof:\\
Let $x$ be arbitrary. [Proof of $P(x)$ goes here]. Since $x$ was arbitrary, we can conclude that $\forall xP(x)$.\\

\noindent Proving a goal of the form $\exists xP(x)$ also involves introducing a variable $x$ but in this case $x$ isn't arbitrary but a value we can choose to bind it with. Since $\exists xP(x)$ is true as long as there's \textit{atleast} one value that makes $P(x)$ true, it's sufficient to prove $P(x)$ is true for just one value.\\

\noindent Strategy:\\
\begin{tikzpicture}
    \beginProofBoxes{
        \textemdash
    }{
        $\exists xP(x)$
    }

    \addProofBoxes{
        $x = $ (the value you decided on)
    }{
        $P(x)$
    }
\end{tikzpicture}\\
Formalized Proof:\\
Let $x = $ (the value you decided on). [Proof of $P(x)$ goes here]. Thus, $\exists xP(x)$.\\

\noindent If you have a given of the form $\exists xP(x)$, you can do \gls{existential instantiation} where you introduce a new variable $x_0$ to stand for an object for which $P(x_0)$ is true, allowing us to assume that $P(x_0)$ is true.\\

\noindent If you have a given of the form $\forall xP(x)$, you can do \gls{universal instantiation} where you plug in any value, say $a$, for $x$ and use this to conclude that $P(a)$ is true (and thus you can treat $P(a)$ as a new given).\\

\noindent Example proof:\\
Suppose $\mathcal{F}$ and $\mathcal{G}$ are families of sets and $\mathcal{F} \cap \mathcal{G} \neq \varnothing$. Prove that $\bigcap \mathcal{F} \subseteq \bigcup \mathcal{G}$. 

\begin{tikzpicture}
    \beginProofBoxes{
        $\mathcal{F} \cap \mathcal{G} \neq \varnothing$
    }{
        $\bigcap \mathcal{F} \subseteq \bigcup \mathcal{G}$
    }

    \addProofBoxes{
        $\exists A(A \in \mathcal{F}, A \in \mathcal{G})$
    }{
        $\forall x(\forall A \in \mathcal{F}(x \in A) \rightarrow \exists A \in \mathcal{G}(x \in A))$
    }

    \addProofBoxes{
        $\exists A(A \in \mathcal{F}, A \in \mathcal{G})$
    }{
        $\forall A \in \mathcal{F}(x \in A) \rightarrow \exists A \in \mathcal{G}(x \in A)$
    }

    \addProofBoxes{
        $\exists A(A \in \mathcal{F}, A \in \mathcal{G})$\\
        $\forall A \in \mathcal{F}(x \in A)$
    }{
        $\exists A \in \mathcal{G}(x \in A)$
    }

    \addProofBoxes{
        $A_0 \in \mathcal{A}$\\ % (existential instantiation)
        $A_0 \in \mathcal{G}$\\ % (existential instantiation)
        $\forall A \in \mathcal{F}(x \in A)$
    }{
        $\exists A \in \mathcal{G}(x \in A)$
    }

    \addProofBoxes{
        $A_0 \in \mathcal{F}$\\
        $A_0 \in \mathcal{G}$\\
        $x \in A_0$\\ % (universal instantiation)
        $\forall A \in \mathcal{F}(x \in A)$
    }{
        $\exists A \in \mathcal{G}(x \in A)$
    }

\end{tikzpicture}\\
On the goals side, we can existentially instantiate $A$ to be $A_0$ (since $A_0 \in \mathcal{G}$), which proves the goal since $x \in A_0$ (note that $x$ is arbitrary).\\

\noindent Another example proof:\\
Suppose $B$ is a set and $\mathcal{F}$ is a family of sets. Prove that if $\bigcup \mathcal{F} \subseteq B$ then $\mathcal{F} \subseteq \mathcal{P}(B)$.\\
\begin{tikzpicture}
    \beginProofBoxes{
        \textemdash
    }{
        $\bigcup \mathcal{F} \subseteq B \rightarrow \mathcal{F} \subseteq \mathcal{P}(B)$
    }

    \addProofBoxes{
        $\bigcup \mathcal{F} \subseteq B$
    }{
        $\mathcal{F} \subseteq \mathcal{P}(B)$
    }

    \addProofBoxes{
        $\bigcup \mathcal{F} \subseteq B$
    }{
        $\forall x(x \in \mathcal{F} \rightarrow x \in \mathcal{P}(B))$
    }

    \addProofBoxes{
        $\bigcup \mathcal{F} \subseteq B$
    }{
        $x \in \mathcal{F} \rightarrow x \in \mathcal{P}(B)$
    }

    \addProofBoxes{
        $\bigcup \mathcal{F} \subseteq B$\\
        $x \in \mathcal{F}$
    }{
        $x \in \mathcal{P}(B)$
    }

    \addProofBoxes{
        $\bigcup \mathcal{F} \subseteq B$\\
        $x \in \mathcal{F}$
    }{
        $x \subseteq B$
    }

    \addProofBoxes{
        $\bigcup \mathcal{F} \subseteq B$\\
        $x \in \mathcal{F}$
    }{
        $\forall y(y \in x \rightarrow y \in B)$
    }

    \addProofBoxes{
        $\bigcup \mathcal{F} \subseteq B$\\
        $x \in \mathcal{F}$
    }{
        $y \in x \rightarrow y \in B$
    }

    \addProofBoxes{
        $\bigcup \mathcal{F} \subseteq B$\\
        $x \in \mathcal{F}$\\
        $y \in x$
    }{
        $y \in B$
    }

    \addProofBoxes{
        $\bigcup \mathcal{F} \subseteq B$\\
        $x \in \mathcal{F}$\\
        $y \in x$
    }{
        $y \in \bigcup \mathcal{F}$
    }

    \addProofBoxes{
        $\bigcup \mathcal{F} \subseteq B$\\
        $x \in \mathcal{F}$\\
        $y \in x$
    }{
        $\exists A \in \mathcal{F}(y \in A)$
    }
\end{tikzpicture}\\
Since $x \in \mathcal{F}$, we can existentially instantiate the set $x$ for $A$ on the goal side. Since $y \in x$, we have just proven the goal.\\

\noindent Theorem again:\\
Suppose $B$ is a set and $\mathcal{F}$ is a family of sets. Prove that if $\bigcup \mathcal{F} \subseteq B$ then $\mathcal{F} \subseteq \mathcal{P}(B)$.\\

\noindent Formalized Proof:\\
Suppose $\bigcup \mathcal{F} \subseteq B$. Let $x$ be an arbitrary element of $\mathcal{F}$. Let $y$ be an arbitrary element of $x$. Since $y \in x$ and $x \in \mathcal{F}$, then by the definition of $\bigcup \mathcal{F}$, $y \in \bigcup \mathcal{F}$. Since $\bigcup \mathcal{F} \subseteq B$, then $y \in B$. Since $y$ was an arbitrary element of $x$, we can conclude that $x \subseteq B$ and thus $x \in \mathcal{P}(B)$. Since $x$ was an arbitrary element of $\mathcal{F}$, we can conclude that $\mathcal{F} \subseteq \mathcal{P}(B)$. Therefore, if $\bigcup \mathcal{F} \subseteq B$ is true, then $\mathcal{F} \subseteq \mathcal{P}(B)$ is true.

\noindent Practice:\\
Theorem: For all integers $a$, $b$, and $c$, if $a \text{ \gls{symb:divides} }b$ and $b \mid c$ then $a \mid c$.\\ 
\begin{tikzpicture}
    \beginProofBoxes{
        \textemdash
    }{
        $\forall a \in \mathbb{Z} \forall b \in \mathbb{Z} \forall c \in \mathbb{Z} (a \mid b \land b \mid c \rightarrow a \mid c)$
    }

    \addProofBoxes{
        $a$, $b$, and $c$ are arbitrary integers
    }{
        $a \mid b \land b \mid c \rightarrow a \mid c$
    }

    \addProofBoxes{
        $a$, $b$, and $c$ are arbitrary integers\\
        $a \mid b$\\
        $b \mid c$
    }{
        $a \mid c$
    }

    \addProofBoxes{
        $a$, $b$, and $c$ are arbitrary integers\\
        $\exists k \in \mathbb{Z} (ka = b)$\\
        $\exists w \in \mathbb{Z} (wb = c)$
    }{
        $\exists q \in \mathbb{Z} (qa = c)$
    }
\end{tikzpicture}\\
Formalized Proof: Let $a$, $b$, and $c$ be arbitrary integers and suppose $a \mid b$ and $b \mid c$. This means there exists some integer $k$ where $ka = b$ is true. This also means there exists some integer $w$ where $wb = c$ is true. This means that $kwa = wb = c$, and since $k$ and $w$ are integers, then the result of multiplying $k$ and $w$ is also an integer. This means there exists some integer when multiplied with $a$ gets $c$. Thus, if $a \mid b$ and $b \mid c$, then $a \mid c$.

\section{Proofs Involving Conjunctions and Biconditionals}
Proving conjunctions and biconditionals is as simple as proving each part individually. So if you have the given $A \land B \land C$, then you just need to prove $A$, $B$, and $C$.\\

\noindent Proving $P \leftrightarrow Q$ (which is equivalent to $P \rightarrow Q \land Q \rightarrow P$) involves proving $P \rightarrow Q$ and proving $Q \rightarrow P$.\\

\noindent Sometimes you can have a string of equivalences to get from $P$ to $Q$ when trying to prove $P \leftrightarrow Q$.\\
Example:\\
Theorem: Suppose $A$, $B$, and $C$ are sets. Then $A \cap (B \setminus C) = (A \cap B) \setminus C$.\\

\noindent Proof:\\
\noindent Let $x$ be arbitrary. Then:
\begin{align}
x \in A \cap (B \setminus C) \textiff& x \in A \land x \in B \setminus C \\
\textiff& x \in A \land x \in B \land x \notin C \\
\textiff& x \in (A \cap B) \land x \notin C \\
\textiff& x \in (A \cap B) \setminus C 
\end{align}
Thus $\forall x(x \in A \cap (B \setminus C) \leftrightarrow x \in (A \cap B) \setminus C)$, so $A \cap (B \setminus C) = (A \cap B) \setminus C$.

\section{Proofs Involving Disjunctions}
When one of your givens is a disjunction, such as $P \lor Q$, there are two possibilities to account for: if $P$ is true, and if $Q$ is true. This means you need to use \glspl{case} where one case is if $P$ is true and the other case is if $Q$ is true and show that both individual possibilities lead to your goal. By considering all possibilities through these cases, this is considered \gls{exhaustive} by Mathematicians. Any proof can be broken up into two or more cases at any time, as long as the cases are exhaustive. Even though cases must be exhaustive, they do not need to be exclusive. Thus, if $P$ and $Q$ happen to both be true, this is okay. Basically: covering some possibilities several times in multiple cases is acceptable as long as all possibilities are covered.\\

\noindent Sometimes a proof has a disjunction for a goal. We can split the proof into cases as long as it's exhaustive. It may be difficult to determine what the cases would be in this scenario. For a goal of $P \lor Q$, you can have case 1 where $P$ is assumed to be true, and in case 2 $P$ is assumed to be false. That means case 1 is trivial and no more work is needed for case 1. For case 2 you need to prove $Q$, because if $\lnot P$ is true, then $Q$ must be true for $P \lor Q$ to be true. Case 2 has a given of $\lnot P$ and a goal of $Q$. Note that this is a simple transformation! Recall that $P \lor Q$ is equivalent to $\lnot P \rightarrow Q$, and then we assume $\lnot P$, leaving $Q$ as the goal. This of course works in reverse, where we assume $\lnot Q$ and try and prove $P$ (achieved via transforming $P \lor Q$ to $\lnot Q \rightarrow P$).\\

\noindent There's a rule of inference called the \gls{disjunctive syllogism}. If you have a given of $P \lor Q$ and also a given of $\lnot P$, then you can conclude that $Q$ is true (adding it to your givens). Similarly, if you have a given of $P \lor Q$ and also a given of $\lnot Q$, then you can conclude that $P$ is true (adding it to your givens).

\clearpage
\printglossary[type=\acronymtype,style=long]  % list of acronyms
\printglossary[type=symbolslist,style=long]   % list of symbols
\printglossary[type=main]                     % main glossary
\end{document}
