\documentclass{article}
\input{preamble}

\title{Relations}
\author{Axel Sorenson}
\date{July 31st, 2024}

\begin{document}
\maketitle

\section{Ordered Pairs and Cartesian Products}
When making a truth set for a statement with more than one free variables, the truth set will not contain individual values, but rather ordered pairs of values. The order of values in an ordered pair makes a difference. That means $(6, 18)$ is not the same as $(18, 6)$. When referring to $a$, $b$, $c$, etc. in an ordered pair/triple/quadruple/etc. $(a, b, c, ...)$, we call them \glspl{coordinate}. $a$ is the first coordinate, $b$ is the second, $c$ is the third, etc.\\

\noindent Suppose $A$ and $B$ are sets. Then the Cartesian product of $A$ and $B$, denoted $A \times B$, is the set of all ordered pairs in which the first coordinate is an element of $A$ and the second is an element of $B$. In other words,
\begin{center}
$A \times B = \{(a,b) \mid a \in A \text{ and } b \in B\}$.
\end{center}

\section{Relations}
Suppose $P(x, y)$ is a statement with two free variables $x$ and $y$. Often such a statement can be thought of as expressing a \gls{relationship} between $x$ and $y$. The truth set of the statement $P(x, y)$ is a set of ordered pairs that records when this relationship holds. In fact, it is often useful to think of any set of ordered pairs in this way, as a record of when some relationship holds. This is the motivation behind the following definition:
\begin{center}
    Suppose $A$ and $B$ are sets. Then a set $R \subseteq A \times B$ is called a \gls{relation} from $A$ to $B$.
\end{center}
The following are definitions regarding relationships:\\
Suppose $R$ is a relation from $A$ to $B$. Then the \gls{domain} of $R$ is the set
\begin{center}
    $\text{Dom}(R) = \{a \in A \mid \exists b \in B((a,b) \in R)\}$.
\end{center}
The \gls{range} of $R$ is the set
\begin{center}
    $\text{Ran}(R) = \{b \in B \mid \exists a \in A((a,b) \in R)\}$.
\end{center}
The \gls{inverse} of $R$ is the set
\begin{center}
    $R^{-1} = \{(b,a) \in B \times A \mid (a,b) \in R\}$.
\end{center}
Finally, suppose $R$ is a relation from $A$ to $B$ and $S$ is a relation from $B$ to $C$. Then the \gls{composition} of $S$ and $R$ is the relation $S \text{ \gls{symb:composition} } R$ from $A$ to $C$ defined as
\begin{center}
    $S \circ R = \{(a,c) \in A \times C \mid \exists b \in B((a,b) \in R \text{ and } (b,c) \in S)\}$.
\end{center}

\bigskip
\noindent The domain of a relation from $A$ to $B$ is the set containing all the first coordinates of ordered pairs in the relation. This in general will be a subset of $A$, but it need not be all of $A$.\\

\noindent The range of a relation from $A$ to $B$ is the set containing all of the second coordinates of ordered pairs in the relation. This in general will be a subset of $B$, but it need not be all of $B$.\\

\noindent The inverse of a relation contains exactly the same ordered pairs as the original relation, but with the order of the coordinates of each pair reversed.\\

\noindent The inverse of the inverse of a relation is equivalent to the relation itself: $(R^{-1})^{-1} = R$.\\ 

\noindent The domain of the inverse of a relation is equivalent to the range of the relation: $\text{Dom}(R^{-1}) = \text{Ran}(R)$. This works in the reverse, where the range of the inverse of a relation is equivalent to the domain of the relation: $\text{Ran}(R^{-1}) = \text{Dom}(R)$.\\

\noindent Compositions are associative, but not commutative. So $T \circ (S \circ R)= (T \circ S) \circ R$ and $A \circ B \neq B \circ A$.

\section{More About Relations}
The notation for a relationship between object $x$ and object $y$ given $R$ is a relation from $A$ to $B$ and $x \in A$ and $y \in B$, $xRy$ means $(x,y) \in R$. This mirrors mathematical relationships such as $x = y$, $x < y$, $x \in y$, $x \subseteq y$, etc. which involves a mathematical symbol between $x$ and $y$ to indicate a relationship between the objects.\\

\noindent A relation $R$ from set $A$ to the same set $A$ is called a \gls{relation on} $A$. For these relations (a relation from a set to the same set), instead of drawing two circles (representing sets), vertices (representing objects within the sets), and edges (representing objects belonging to ordered pairs in the relation), Mathematicians often represent these relations with \glspl{directed graph}.\\

\noindent Some definitions:\\
Suppose $R$ is a relation on $A$.
\begin{enumerate}
    \item $R$ is said to be \gls{reflexive} on $A$ (or just reflexive, if $A$ is clear from context) if $\forall x \in A(xRx)$, or in other words $\forall x \in A((x,x) \in R)$.
    \item $R$ is said to be \gls{symmetric} if $\forall x \in A \forall y \in A(xRy \rightarrow yRx)$.
    \item $R$ is said to be \gls{transitive} if $\forall x \in A \forall y \in A \forall z \in A((xRy \land yRz) \rightarrow xRz)$.
    \item $R$ is said to be \gls{antisymmetric} if $\forall x \in A \forall y \in A((xRy \land yRx) \rightarrow x = y)$
\end{enumerate}
If a relation is reflexive, then its directed graph will have \glspl{loop} on all its vertices.\\

\noindent Suppose $A$ is a set. The \gls{identity relation} $i_{A} = \{(x,y) \in A \times A \mid x = y \}$. The identity relation could also be defined as $i_{A} = \{(x,x) \mid x \in A \}$. The identity relation $i_{A}$ for any set $A$ will be reflexive, symmetric, and transitive.\\

\noindent The following connects the properties and operations defined about sets so far:\\
Suppose $R$ is a relation on $A$.
\begin{enumerate}
    \item $R$ is reflexive$\textiff i_{A} \subseteq R$, where as before $i_{A}$ is the identity relation on $A$.
    \item $R$ is symmetric$\textiff R = R^{-1}$.
    \item $R$ is transitive$\textiff R \circ R \subseteq R$.
\end{enumerate}

\section{Ordering Relations}
Suppose $R$ is a relation on a set $A$. Then $R$ is called a \gls{partial order} on $A$ (or just a partial order if $A$ is clear from context) if it is reflexive, transitive, and antisymmetric. It is called a \gls{total order} on $A$ (or just a total order if $A$ is clear from context) if it is a partial order, and in addition it has the following propery:
\begin{center}
    $\forall x \in A \forall y \in A(xRy \lor yRx)$.
\end{center}


\clearpage
\printglossary[type=\acronymtype,style=long]  % list of acronyms
\printglossary[type=symbolslist,style=long]   % list of symbols
\printglossary[type=main]                     % main glossary
\end{document}
